Neural Circuit Organization of the Ctenophore Apical Organ

### 2025-03-18 manuscript/Manuscript.qmd
edit figure legends

---
title: "Neural Circuit Organization of the Ctenophore Apical Organ and Its Role in Gravity Sensing"

@@

![**Figure 1. Morphology of the apical organ in *Mnemiopsis leidyi***\
(A) Whole-body image of a 5-day-old M. leidyi cydippid larva in lateral view, observing the tentacular plane. The boxed region indicates the apical organ. Scale bar: 100 µm.\
(B) Enlarged views of the apical organ from different perspectives in a 5-day-old M. leidyi cydippid larva. The left panel shows the apical organ from a aboral view. The middle panel presents a sagittal view of the apical organ. The right panel provides a tentacular plane view of the apical organ. Scale bar: 10 µm.\
(C) Schematic diagram of the serial sectioning process. The entire cydippid larva was embedded in resin, trimmed around the apical organ, and sectioned into serial ultrathin sections, which were collected on glass slides in a ribbon-like arrangement.\
(D) Example of cell tracing using the collaborative annotation toolkit CATMAID for large-scale electron microscopy image datasets. The spherical objects indicate nuclear positions, while the lines represent the traced cell centers. Scale bar: 10 µm.\
(E) Three-dimensional reconstruction of all cells composing the apical organ, displayed from different perspectives. The left panel shows a aboral view of the apical organ. The middle panel presents a sagittal plane view. The right panel provides a tentacular plane view. Cells are color-coded according to their types. Scale bar: 10 µm.](figures/Figure1.png)

@@

![**Figure 2. Reconstruction of Three Syncytial Neurons in the Apical Organ**\
(A) 3D reconstruction of the aboral nerve net (ANN) Q1Q2Q3Q4 (Q1-4), which spans all four quadrants and contains multiple nuclei (light blue). The spheres represent the positions of individual nuclei. The left panel shows a aboral view of the apical organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(B) 3D reconstruction of two ANN Q1Q2 (pink) and Q3Q4 (orange), each spanning two quadrants and containing multiple nuclei. The spheres indicate the positions of the nuclei. The left panel shows a aboral view of the apical organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(C) The abbreviations are as follows: bal (balancer), brg (bridge), bsl (bristle), cg (ciliated groove), dv (dense vesicle cells), imc (intra-multiciliated cells), la (lamellate bodies), li (lithocytes), pl (plumose), and ef (epithelial floor cells).
(D) Localization of mitochondria within AO neuron Q1Q2Q3Q4. Red indicates mitochondria associated with the presynaptic triad structures, yellow marks mitochondria containing synaptic vesicles but lacking a clearly defined presynaptic triad, blue represents mitochondria with unclear synaptic vesicles, and black denotes mitochondria where no synaptic vesicles were identified.\
(A) Representative electron micrographs of presynaptic triad structures within the dataset. The positions of mitochondria are marked in orange. Postsynaptic target cells were inferred based on the locations of synaptic vesicles and the endoplasmic reticulum, with light blue arrows indicating the direction across the cell membrane. The left panel illustrates a monadic structure, while the right panel shows a polyadic (or dyadic) structure. Scale bar: 500 nm.](figures/Figure2.png)

@@

Each cell is assigned a specific type name based on its category (balancer, biC, bridge, bristle, dense vesicle cell, dome, EF, IcMC, LB, lithocyte, monoC, non-C, plumose cell, SNN, ssCG, SSN, stCG), resulting in a total of 17 cell types. Additionally, a quadrant number is appended to the cell type name following an underscore ("\_") to indicate the cell body’s location. For example, cells located in the first quadrant are annotated with "Q1." If a cell is situated between the first and second quadrants, the annotation "Q1Q2" is applied.

>>>>

---
title: "Neural Circuit Organization of the Ctenophore Apical Organ"

@@

![**Figure 1. Morphology of the apical organ in *Mnemiopsis leidyi***\
(A) Whole-body image of a 5-day-old *M. leidyi* cydippid larva in lateral view. The boxed region indicates the apical organ (ao). Scale bar: 100 µm.\
(B) Schematic diagram of the aboral view of a cydippid larva. The two-tone coloration represents the biradial symmetry of the body. For convenience, the body is divided into four quadrants, designated as the first (Q1) through fourth (Q4) quadrants. The two primary viewing angles are referred to as the sagittal plane (S) and the tentacular plane (T). Abbreviations: ao, apical organ; cg, ciliated groove; cr, comb plates; tb, tentacle bulbs.\
(C) Enlarged views of the apical organ from different perspectives in a cydippid larva using a DIC microscope. The left panel presents an aboral view of the apical organ. The middle panel shows a lateral view of the apical organ in the sagittal (S) plane, while the right panel displays a lateral view in the tentacular (T) plane. Abbreviations: bal, balancer; cg, ciliated grooves; do, dome cilia; li, lithocyte. Scale bar: 50 µm.\
(D) Schematic representations of the apical organ from three different perspectives. The left panel illustrates the aboral view, the middle panel shows the lateral view in the sagittal plane, and the right panel presents the lateral view in the tentacular plane. The quadrant colors correspond to those in panel (B).\
(E) Example of cell tracing using the collaborative annotation toolkit CATMAID for large-scale electron microscopy image datasets. The spherical objects indicate nuclear positions, while the lines represent the traced cell centers. Scale bar: 10 µm.\
(F) Three-dimensional reconstruction of cells composing the apical organ, displayed from different perspectives. The left panel presents an aboral view of the apical organ. The middle panel shows a sagittal plane view, while the right panel provides a tentacular plane view. Cells are color-coded according to their respective quadrants. Lithocytes (li) are represented as three gray spheres, while balancers (bal) are depicted as gray lines. Scale bar: 25 µm.](figures/Figure1.png)

@@

![**Figure 2. Identification of Three Syncytial Neurons in the Apical Organ**\
(A) 3D reconstruction of the aboral nerve net (ANN) Q1Q2Q3Q4 (Q1-4), which spans all four quadrants and contains multiple nuclei (blue). The spheres represent the positions of individual nuclei. The left panel shows a aboral view of the apical organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(B) 3D reconstruction of two ANN Q1Q2 (pink) and Q3Q4 (orange), each spanning two quadrants and containing multiple nuclei. The spheres indicate the positions of the nuclei. The left panel shows a aboral view of the apical organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(C) Graph showing the number of mitochondria per cell and the proportion of mitochondria involved in synapse formation for each cell type. Green represents mitochondria associated with synapses, while gray represents mitochondria not involved in synapse formation. ANN has a significantly higher number of mitochondria compared to other cell types, with most of them forming synapses. The abbreviations are as follows: bal (balancer), brg (bridge), bsl (bristle), cg (ciliated groove), dv (dense vesicle cells), imc (intra-multiciliated cells), la (lamellate bodies), li (lithocytes), pl (plumose), and ef (epithelial floor cells).\
(D) Localization of mitochondria within AO neuron Q1Q2Q3Q4. Red indicates mitochondria associated with the presynaptic triad structures, yellow marks mitochondria containing synaptic vesicles but lacking a clearly defined presynaptic triad, blue represents mitochondria with unclear synaptic vesicles, and black denotes mitochondria where no synaptic vesicles were identified.\
(E) Representative electron micrographs of presynaptic triad structures observed in the dataset and their schematic diagrams. The left diagram illustrates synaptic projections from the center ANN (ANN Q1-4) to the lateral ANN (ANN Q3Q4), while the right diagram shows synaptic projections from the lateral ANN (ANN Q1Q2) to the center ANN (ANN Q1-4). The abbreviations are as follows: mi (mitochondrion), er (endoplasmic reticulum), and sv (synaptic vesicles).\
(F) Plot showing the three-dimensional positions of synapses. Synapses from the center ANN to the lateral ANN are shown in magenta, while synapses from the lateral ANN to the center ANN are shown in cyan. Blue dots indicate the locations of autapses within ANN Q1-4.\
(G) Neural circuit map of ANN connectivity. The blue, orange, and pink circles represent individual neurons in ANN Q1-4, Q2Q3, and Q3Q4, respectively. Magenta arrows indicate synaptic connections from the center ANN to the lateral ANN, while cyan arrows indicate connections from the lateral ANN to the center ANN. Numbers denote the number of synapses. Blue arrows represent autapses in ANN Q1-4 along with their synapse counts.](figures/Figure2.png)

@@

Each cell is assigned a specific type name based on its category (balancer, biC, bridge, bristle, dense vesicle cell, dome, EF, IcMC, LB, lithocyte, monoC, non-C, plumose cell, SNN, ssCG, ANN, stCG), resulting in a total of 17 cell types. Additionally, a quadrant number is appended to the cell type name following an underscore ("\_") to indicate the cell body’s location. For example, cells located in the first quadrant are annotated with "Q1." If a cell is situated between the first and second quadrants, the annotation "Q1Q2" is applied.
